---
title: "Speed Dating Dynamics: An Investigation by Group 6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

#### Speed Dating Dynamics: An Investigation by Group 6

In this study, we delve into the intriguing world of speed dating, utilizing a comprehensive dataset that captures a myriad of interactions, preferences, and decisions made during speed dating events. The dataset, titled "SpeedDating.csv", is a rich compilation of demographic information, self-reported attributes, and individual preferences of participants engaged in speed dating. It includes data points such as age, gender, interests, and crucially, participants' ratings of their dates across various attributes like attractiveness, sincerity, intelligence, fun, and ambition. Each entry also records the critical decision of whether a participant would like to see their date again, offering a direct insight into dating preferences and choices.

The central aim of our research is to unravel the attributes that significantly influence the decision to date a partner in a speed dating context. Specifically, we seek to understand how attributes like sincerity, fun, intelligence, and notably, physical attractiveness, play a role in the decision-making process of both genders. Our analysis is poised to examine the relationships between gender-specific ratings of these attributes and the likelihood of wanting to pursue a further date.

We anticipate uncovering distinct trends and patterns that reveal gender differences in dating preferences. Preliminary observations suggest that physical attractiveness may be a more critical factor for men than women in deciding whether to date someone. In contrast, we expect to find that women may place relatively higher importance on other attributes like intelligence or sincerity. This investigation aims not only to present a descriptive analysis of the speed dating dynamics but also to offer deeper insights into the underlying factors that drive dating choices in a fast-paced and increasingly popular dating environment.

By applying a range of statistical methods and data visualization techniques, we aim to present a nuanced and comprehensive analysis of speed dating dynamics. Our approach is grounded in a robust statistical framework, leveraging the R programming environment to manipulate, analyze, and visualize the data. Through this exploration, we hope to contribute meaningful insights to the understanding of modern dating behaviors and preferences.

# Methods

To adress our three research questions 

<div class="container">

```{r setup, echo=FALSE}
library(tidyverse)
library(dplyr)
library(stringr)
```



```{r load file, echo=FALSE}
load("SpeedDating.rda")

head(SpeedDating)
```
Question: What attributes (like sincerity, fun, intelligence) most significantly influence the decision to date a partner?

To answer this question we wanted to examine the relationships that existed, if any, between each gender's rating of their dates' attributes and the chances of saying "yes" to wanting another date. Here is an analyses of the results from when we used data visualization to answer the question in regard to the physical attractiveness trait for each person. 

```{r AttractiveM, echo=FALSE}

SpeedDating %>% 
  ggplot(mapping = aes(x = AttractiveM, fill = factor(DecisionM))) +
  geom_bar() +
  labs(x = "Male rating of Female's attractiveness", 
       y = "Count",
       title = "Male's Decision based on Female Attractiveness") +
  scale_fill_discrete(name = "Does the Male want \nanother date?", labels = c("No", "Yes"))

```

There are some very interesting conclusions that we can draw from this graph. First, we see the majority of "yes" answers from the men fall around a 7-8 on the physical attractiveness scale of how he would rate the woman. We can also see that no man said he wanted to have another date with a woman that he rated below a 5. Both of these observations suggest that physical attractiveness is an important deciding factor to men when choosing a partner.


```{r AttractiveF, echo=FALSE}

SpeedDating %>% 
  ggplot(mapping = aes(x = AttractiveF, fill = factor(DecisionF))) +
  geom_bar() +
  labs(x = "Female rating of Male's attractiveness",
       y = "Count",
       title = "Female's Decision based on Male Attractiveness") +
  scale_fill_discrete(name = "Does the Female want \nanother date?", labels = c("No", "Yes"))

```
We now see very different behavior in the women's decisions compared to the men's. Notably, some women said "yes" to another date for a man that they rated at the lowest end of the physical attractiveness scale, *and* more women said "no" to men that they would rate around a 7-8 than the men said "no" to women at this level off attraction. These observations imply that women do not consider physical attractiveness to be the most important factor when choosing a partner.


In conclusion, the data shows that men value physical attractiveness more than women do in a potential partner. However, we should keep in mind that the data was collected from a group of mainly young adults in their twenties and thirties, so this conclusion may not be true for the general population. Nonetheless, this was a very interesting data set to explore for this project.  



```{r setup, echo=FALSE}
library(tidyverse)
library(dplyr)
library(stringr)

```



```{r load file, echo=FALSE}
load("SpeedDating.rda")

head(SpeedDating)
```
Question: What attributes (like sincerity, fun, intelligence) most significantly influence the decision to date a partner?

To answer this question we wanted to examine the relationships that existed, if any, between each gender's rating of their dates' attributes and the chances of saying "yes" to wanting another date. Here is an analyses of the results from when we used data visualization to answer the question in regard to the physical attractiveness trait for each person. 

```{r AttractiveM, echo=FALSE}

SpeedDating %>% 
  ggplot(mapping = aes(x = AttractiveM, fill = factor(DecisionM))) +
  geom_bar() +
  labs(x = "Male rating of Female's attractiveness", 
       y = "Count",
       title = "Male's Decision based on Female Attractiveness") +
  scale_fill_discrete(name = "Does the Male want \nanother date?", labels = c("No", "Yes"))

```

There are some very interesting conclusions that we can draw from this graph. First, we see the majority of "yes" answers from the men fall around a 7-8 on the physical attractiveness scale of how he would rate the woman. We can also see that no man said he wanted to have another date with a woman that he rated below a 5. Both of these observations suggest that physical attractiveness is an important deciding factor to men when choosing a partner.


```{r AttractiveF, echo=FALSE}

SpeedDating %>% 
  ggplot(mapping = aes(x = AttractiveF, fill = factor(DecisionF))) +
  geom_bar() +
  labs(x = "Female rating of Male's attractiveness",
       y = "Count",
       title = "Female's Decision based on Male Attractiveness") +
  scale_fill_discrete(name = "Does the Female want \nanother date?", labels = c("No", "Yes"))

```
We now see very different behavior in the women's decisions compared to the men's. Notably, some women said "yes" to another date for a man that they rated at the lowest end of the physical attractiveness scale, *and* more women said "no" to men that they would rate around a 7-8 than the men said "no" to women at this level off attraction. These observations imply that women do not consider physical attractiveness to be the most important factor when choosing a partner.


In conclusion, the data shows that men value physical attractiveness more than women do in a potential partner. However, we should keep in mind that the data was collected from a group of mainly young adults in their twenties and thirties, so this conclusion may not be true for the general population. Nonetheless, this was a very interesting data set to explore for this project.  

```{r cars}
sd <- read.csv("SpeedDating.csv", header = TRUE)
sd <- as_tibble(sd)
glimpse(sd)
```

```{r}
as.double(sd$SincereM)
glimpse(sd)
```

```{r}
New_sd <- sd%>%
  filter(!is.na(AttractiveM))%>%
  filter(!is.na(AttractiveF))%>%
  filter(!is.na(IntelligentM))%>%
  filter(!is.na(IntelligentF))%>%
  filter(!is.na(FunM))%>%
  filter(!is.na(FunF))%>%
  filter(!is.na(AmbitiousM))%>%
  filter(!is.na(AmbitiousF))%>%
  filter(!is.na(SharedInterestsM))%>%
  filter(!is.na(SharedInterestsF))%>%
  filter(!is.na(SincereM))%>%
  filter(!is.na(SincereF))%>%
  summarise(Avg_attractiveM = mean(AttractiveM),
            Avg_attractiveF= mean(AttractiveF),
            Avg_SincereM = mean(SincereM),
            Avg_SincereF = mean(SincereF),
            Avg_IntelligentM = mean(IntelligentM),
            Avg_IntelligentF = mean(IntelligentF),
            Avg_FunM = mean(FunM),
            Avg_FunF = mean(FunF),
            Avg_AmbitiousM = mean(AmbitiousM),
            Avg_AmbitiousF = mean(AmbitiousF),
            Avg_SharedInterestsM = mean(SharedInterestsM),
            Avg_SharedInterestsF = mean(SharedInterestsF))
New_sd
```
```{r}
sd %>% 
  ggplot(mapping = aes(x = AttractiveM, fill = factor(DecisionM))) +
  geom_bar() +
  labs(x = "Male rating of Female's attractiveness", 
       title = "Male's Decision based on Female Attractiveness") +
  scale_fill_discrete(name = "Yes or no", labels = c("No", "Yes"))
```

```{r}
sd %>% 
  ggplot(mapping = aes(x = AttractiveF, fill = factor(DecisionF))) +
  geom_bar() +
  labs(x = "Female rating of Male's attractiveness", 
       title = "Female's Decision based on Male Attractiveness") +
  scale_fill_discrete(name = "Yes or no", labels = c("No", "Yes"))
```

```{r}
sd %>% 
  ggplot(mapping = aes(x =IntelligentM , fill = factor(DecisionM))) +
  geom_bar() +
  labs(x = "Male rating of Female Intelligence", 
       title = "Male's Decision based on Female Intelligence")+
  scale_fill_discrete(name = "Yes or no", labels = c("No", "Yes"))
```
```{r}
sd %>% 
  ggplot(mapping = aes(x = IntelligentF, fill = factor(DecisionF))) +
  geom_bar() +
  labs(x = "Female rating of Male's Intellingence", 
       title = "Female's Decision based on Male Intelligence") +
  scale_fill_discrete(name = "Yes or no", labels = c("No", "Yes"))
```


```{r}
sd %>% 
  ggplot(mapping = aes(x =FunM , fill = factor(DecisionM))) +
  geom_bar() +
  labs(x = "Male rating on how fun female is", 
       title = "Male's Decision based on How Fun Female is") +
  scale_fill_discrete(name = "Yes or no", labels = c("No", "Yes"))
```
```{r}
sd %>% 
  ggplot(mapping = aes(x = FunF, fill = factor(DecisionF))) +
  geom_bar() +
  labs(x = "Female rating on how fun Male is", 
       title = "Female's Decision based on how fun Male is") +
  scale_fill_discrete(name = "Yes or no", labels = c("No", "Yes"))
```



```{r}
sd %>% 
  ggplot(mapping = aes(x = SincereM, fill = factor(DecisionM))) +
  geom_bar() +
  labs(x = "Male rating on how Sincere female is", 
       title = "Male's Decision based on how Sincere female is") +
  scale_fill_discrete(name = "Yes or no", labels = c("No", "Yes"))
```

```{r}
sd %>% 
  ggplot(mapping = aes(x = SincereF, fill = factor(DecisionF))) +
  geom_bar() +
  labs(x = "Female rating on how Sincere Male is", 
       title = "Female's Decision based on how Sincere  is") +
  scale_fill_discrete(name = "Yes or no", labels = c("No", "Yes"))
```



```{r}
sd %>% 
  ggplot(mapping = aes(x =AmbitiousM , fill = factor(DecisionM))) +
  geom_bar() +
  labs(x = "Male rating on how Ambitious female is", 
       title = "Male's Decision based on how Ambitious female is") +
  scale_fill_discrete(name = "Yes or no", labels = c("No", "Yes"))
```

```{r}
sd %>% 
  ggplot(mapping = aes(x =AmbitiousF , fill = factor(DecisionF))) +
  geom_bar() +
  labs(x = "Female rating on how Ambitious Male is", 
       title = "Female's Decision based on how Ambitious Male is") +
  scale_fill_discrete(name = "Yes or no", labels = c("No", "Yes"))
```


```{r}
sd %>% 
  ggplot(mapping = aes(x =SharedInterestsM , fill = factor(DecisionM))) +
  geom_bar() +
  labs(x = "Male rating on Shared interest with Females", 
       title = "Male's Decision based on Shared interest with Females") +
  scale_fill_discrete(name = "Yes or no", labels = c("No", "Yes"))
```

```{r}
sd %>% 
  ggplot(mapping = aes(x =SharedInterestsF, fill = factor(DecisionF))) +
  geom_bar() +
  labs(x = "Female rating on Shared interest with Male", 
       title = "Female's Decision based on Shared interest with Male") +
  scale_fill_discrete(name = "Yes or no", labels = c("No", "Yes"))
```


Q. Can we predict the likelihood of a successful match based on the participants' ratings and demographics?
```{r}
library(tidyverse)
library(caret)
speed_dating_data <- read.csv("SpeedDating.csv")
speed_dating_data <- speed_dating_data %>%
  mutate_if(is.numeric, ~ifelse(is.na(.), median(., na.rm = TRUE), .)) %>%
  mutate_if(is.character, ~ifelse(is.na(.), Mode(.), .)) # Mode function needed
speed_dating_data$RaceM <- as.factor(speed_dating_data$RaceM)
speed_dating_data$RaceF <- as.factor(speed_dating_data$RaceF)
speed_dating_data$Match <- ifelse(speed_dating_data$DecisionM == 1 & speed_dating_data$DecisionF == 1, 1, 0)
speed_dating_data <- speed_dating_data %>% select(-DecisionM, -DecisionF)
model_data <- model.matrix(~ . - 1, data = speed_dating_data)
model_data <- as.data.frame(model_data)
set.seed(42)
training_rows <- createDataPartition(model_data$Match, p = 0.7, list = FALSE)
train_data <- model_data[training_rows, ]
test_data <- model_data[-training_rows, ]
model <- glm(Match ~ ., data = train_data, family = "binomial")
predictions <- predict(model, newdata = test_data, type = "response")
predicted_classes <- ifelse(predictions > 0.5, 1, 0)
confusionMatrix(factor(predicted_classes), factor(test_data$Match))

```

```{r}
library(ggplot2)

conf_matrix <- confusionMatrix(factor(predicted_classes), factor(test_data$Match))
conf_matrix_df <- as.data.frame(conf_matrix$table)

ggplot(conf_matrix_df, aes(x = Reference, y = Prediction)) +
  geom_tile(aes(fill = Freq), color = "white") +
  geom_text(aes(label = Freq), vjust = 1) +
  scale_fill_gradient(low = "lightblue", high = "blue") +
  theme_minimal() +
  labs(title = "Confusion Matrix", x = "Actual Label", y = "Predicted Label") +
  theme(plot.title = element_text(hjust = 0.5))


```
ROC Curve and AUC
```{r}
library(pROC)

roc_obj <- roc(test_data$Match, predictions)
auc(roc_obj)

plot(roc_obj, main = "ROC Curve with AUC")
abline(a = 0, b = 1, col = "red")
text(0.6, 0.2, paste("AUC =", round(auc(roc_obj), 2)))


```
Coefficient Importance Plot

```{r}
library(ggplot2)
coefficients <- as.data.frame(coef(model))
names(coefficients) <- c("Estimate")
coefficients$Feature <- rownames(coefficients)
rownames(coefficients) <- NULL
ggplot(coefficients, aes(x = Feature, y = Estimate)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  coord_flip() +
  labs(title = "Feature Coefficients", x = "Feature", y = "Coefficient")

```

